### 2017-12-18 paper/introduction.tex
backup

The remainder of this paper is organized as follows:
Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 details the various architecture optimizations we used in FlexGraph, section 5 describes FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 6 describes the experimental setup, section 7 describes our results analysis, section 8 describes the related work, section 9 summarizes our main contribution and results.

>>>>

The remainder of this paper is organized as follows:
Section 2 provides a background and motivation for the FlexGrah accelerator, section 3 describes FlexGraph's architecture, section 4 describes FlexGraph's software-hardware codesign using Cash \cite{Cash} framework, section 5 describes the experimental setup, section 6 describes our results analysis, section 7 describes the related work, section 8 summarizes our main contribution and results.

### 2017-12-18 paper/mechanism.tex
backup

\begin{figure}[ htbp ] 
\begin{equation}
\#Edges = A^{T} \times Identity
\end{equation}
\begin{equation}
\#Edges = 
\left(\begin{array}{ccc} 0 & 0 & 0 \\ 1 & 0 & 0 \\ 1 & 1 & 0 \end{array}\right) \times
\left(\begin{array}{c} 1 \\ 1 \\ 1 \end{array}\right) 
= \left(\begin{array}{c} 0 \\ 1 \\ 2 \end{array}\right) 
\end{equation}
\caption{Incoming Edges Algorithm}
\label{fig:sample_algorithm}
\end{figure}

\begin{lstlisting}[emph={Assign, Process, Apply}, emphstyle=\color{blue}, numbers=none, escapechar=\%, caption=Graph Vertex Processing Model]
GraphProgram(%\textit{G}%, %\textit{P}%, %\textit{N}%) :
  for_each %\textit{N}% :
    x := Assign(%\textit{P}%)
  	y := Process(%\textit{G}%,x)
  	%\textit{P}% := Apply(%\textit{P}%,y)

@@

Listing 1 shows the three stages of a generalized graph vertex processing model - \textit{Assign}, \textit{Process} and \textit{Apply}, given a graph \textit{G}, some vertex properties \textit{P} and an iteration count \textit{N}. The \textit{Assign} stage generates an input vector \textit{x} using the vertex properties \textit{P}. The \textit{Process} stage applies the input vector \textit{x} to the graph \textit{G} and generates an output vector \textit{y}. The\textit{ Apply} gather the resulting output vector to update the vertex properties for the next iteration. The graph program executes the three stages for several iterations until it converges. A direct mapping of our incoming edges algorithm to this processing model is also provided in Listing 1, where the identity vector is used for the vector properties. The \textit{Assign} and \textit{Apply} stages of this program are simple identity operators, while the \textit{Process} stage perform a matrix-vector multiplication. The input graph can be represented as a compressed sparse matrix to only store the non-zero entries of the graph. This data structure enables accelerated computation, often using a Sparse Matrix-Sparse Vector Multiplication (SpMSpV) kernel to target multi-core CPUs \cite{GraphMat} or GPUs \cite{MapGraph}. The computation that cannot be accelerated is simply executed on the general-purpose host CPU. FlexGraph's objective is to provide a more energy efficient graph processing backend using a FPGA for hardware acceleration \cite{Catapult}.

@@

\begin{figure*}
\centering
\begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figures/offload_model}
    \caption{Offload model}
    \label{fig:offload_model}
\end{subfigure}
\hfill
\begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figures/collaborative_model}
    \caption{Collaborative model}
    \label{fig:collaborative_model}
\end{subfigure}
\caption{Shared Memory Compute Models}
\label{fig:compute_model}
\end{figure*}

@@

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\textwidth]{figures/DCSC_matrix_format}
\caption{DCSC hyperspace matrix format}
\label{fig:DCSC_matrix_format}
\end{figure}

FlexGraph uses the Doubly Compressed Sparse Column (DCSC) \cite{DCSC} format as underlying graph data structure. The format allows minimal traversal into the sparse matrix structure, saving necessary memory bandwidth when fetching empty columns. Figure \ref{fig:DCSC_matrix_format} shows a sample matrix A = \{(0,5,a), (0,7,b), (6,3,c), (7,1,d)\} with four non-zero edges (src, dst, weight) encoded using DCSC versus the conventional CSC \cite{CSC} format. The traversal over CSC requires accessing all consecutive column ranges \textit{(cols[i+2] - cols[i])} even though most of the distances are empty. DCSC saves on bandwidth by using an additional indirection  buffer inside its structure to only encode non-zero columns. It is important to note that this indirection can introduce additional storage overhead and access latency if the matrix is too small or not sparse enough. However, Graph Analytics datasets are generally very large and hyper-sparse which eliminate the DCSC format overhead. In addition to a sparse matrix, FlexGraph also uses a sparse vector to provide property data as well as edge selection during computation. A bitmask is used to encode the non-zero entries in the input vector.

@@

FlexGraph micro-architecture implements a Sparse Matrix-Sparse Vertex Multiplication (SpMSpV) kernel in FPGA. This kernel diverges from conventional Sparse Matrix-Vector Multiplication (SpMV) implementations in two ways - firstly, it uses a sparse data structure as input and output vector during computation - secondly, it uses a doubly-compressed sparse matrix (DCSC) format with additional memory indirection buffer when accessing the matrix columns. These two properties present unique performance challenges when designing the accelerator. Listing 1 shows the pseudo-code of the SpMSpV kernel. Given the start/end addresses (col\_start, col\_end) into the matrix columns buffer (coldata), for The program iterates through each column entry and fetches the corresponding column index (col) and rows buffer address (row\_start, row\_end). It then checks if the current column index (col) is active using the vertex bitmask, only then it proceeds with rows traversal loop where the matrix row\_data is accessed to retrieve corresponding row index and matrix value for computation. A Multiply-Accumulate (MAC) operation is then performed on the matrix and vertex data and output bitmask is updated. Lines 6 and 8 show the code region where external memory is randomly accessed to obtain the vertex data and active state. Lines 5 and 10 show the code region where external memory is accessed semi-random because only the first access cannot be predicted and after that the address is simply incremented. FlexGraph's architecture attempts to address some of those performance hogs using several optimizations detailed in section 4.

\begin{lstlisting}[
	linebackgroundcolor={
	\ifnum\value{lstnumber}=6\color{red!10}\fi
	\ifnum\value{lstnumber}=8\color{red!10}\fi
	\ifnum\value{lstnumber}=5\color{orange!10}\fi
	\ifnum\value{lstnumber}=10\color{orange!10}\fi}, 
        escapechar=\%, 
        emph={col_start, col_end, coldata, rowdata, values, active}, emphstyle=\color{blue}, language=Python, caption=Pseudo-code for SpMSpV kernel]
def SpMSpV_kernel(a, x):
  y_values[] = {0}
  y_bitmask[] = {false}
  for i in (a.col_start, a.col_end):            	
    (col, row_start, row_end) = a.coldata[i]
    x_active = x.bitmask[col]          
    if (x_active) :                            
      x_value  = x.values[col]              
      for j in (row_start, row_end): 
        (row, a_value) = a.rowdata[j]
        y_values[row] += a_value * x_value
        y_bitmask[row] = true          
  return (y_values, y_bitmask)
\end{lstlisting}

@@

Externally, FlexGraph input and output signals implement an Accelerator Functional Unit (AFU) interface defined by Intel's Accelerator Abstraction Layer (AAL) \cite{Intel-FPGA}. AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL software running on the host processor. Listing 2 shows AAL device interface implementation using Cocoh's API. Our FlexGraph Accelerator's class in Cocoh simply derives from this interface and override the \textit{initialize()} function to provide its implementation and Cocoh takes care of generating the corresponding verilog module. It implements three input signals \textit{(start, qpi\_in, ctx)} and two output signals \textit{(qpi\_out, done)}. The AFU starts execution when the \textit{start} signal is asserted and communicate completion by asserting the \textit{done} signal. The \textit{ctx} signal provides application's specific data like constants and buffers address in the case of FlexGraph. The \textit{qpi} in/out signals implement Intel Quick Path Interconnect (QPI) interface \cite{QPI}, providing single channel read/write ports for accessing external shared memory. The FPGA socket hosts a 64-byte cache coherent interface (CCI) \cite{CCI} that connects to the host processor's last level cache (LLC) via QPI.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.4\textwidth]{figures/microarchitecture}
\caption{FlexGraph Microarchitecture}
\label{fig:microarchitecture}
\end{figure}

Figure \ref{fig:microarchitecture} illustrates FlexGraph accelerator microarchitecture. It is comprised of four main module types, a controller, processing elements (PEs), a load store unit (LSU) and vertex caches. The main controller is responsible for starting the accelerator, scheduling tasks for execution, reporting hardware counters and terminating the execution. The processing elements execute partition tasks assigned to them by the controller and communicate with the LSU to access the matrix and vertex data for their partition. They are also responsible for sending their final output result back to memory. The LSU is the module responsible for managing external communication between the accelerator and memory via the QPI interface. It directly binds to the QPI ports defined in listing 2.
The vertex caches store intermediate vertex data and active masks for sharing between the processing elements. FlexGraph processing pipeline sightly resembles the SPMV execution steps illustrated in listing 1. We made several important modifications to it to improve performance.

\section{FlexGraph Optimizations}

The first optimization we performed in FlexGraph early during the design phase was the support of multiple processing units. The reasoning behind it was to extract the maximum bandwidth out of the LSU such that the QPi is always busy processing a request when some processing element are stalled waiting for their data to return. the other advantage of parallelization for FlexGraph is to increase the overall accelerator throughput by processing multiple a larger chunk of the workload per unit of time. The scheduling of the partitions for execution on the processing elements is controlled by a dispatch unit inside the main controller module. The dispatch unit fetches partitions data from the LSU and pass down each partition in a first come first serve fashion to the processing elements.

\subsection{DCSC Matrix partitioning}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.4\textwidth]{figures/matrix_partitioning}
\caption{DCSC Matrix Partitioning}
\label{fig:matrix_partitioning}
\end{figure}

To enable the parallelization of FlexGraph tasks, the DCSC sparse matrix is first partitioned into aligned partitions of 32 consecutive rows containing non-zero edges. we choose a partition size of 32 mainly to match the 32-bit size of the bitmasks encoding the vertices that are active. These masks are used by the software on the host processor to determine the regions of the acceleration's output buffer that have been updated. Figure \ref{fig:matrix_partitioning} illustrates sample matrix partitioning in which the non-zero horizontal regions have been broken into two partitions P0 and P1. The non-zero column ranges \textit{(c0, c1, c2, c3)} represent the selected chunks covered by each partition. Partition P0 will only contain ranges \textit{c0} and \textit{c2} while P2 has non-zeros in all four ranges. The partitioning scheme is not ideal because of the workload in-balance that might exist when the number of non-zero varies disproportionally between partitions.

\subsection{Synchronising Memory Accesses}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.35\textwidth]{figures/write_masks_synchronization}
\caption{Active Masks Write Synchronization}
\label{fig:write_masks_synchronization}
\end{figure}

An implementation challenge we faced when supporting multiple processing elements in FlexGraph was the synchronization of write accesses for the output active masks. Because the LSU data transfer granularity is 64 byte (matching CCI \cite{CCI}), FlexGraph need to manage shared write accesses to the same block to avoid having a processing element override the content of another one. Luckily this doesn't pose any problem for writing out partition output values because the 32 rows within a partition occupy 32 * 4 bytes = 2 x 64 bytes blocks. Because the partitioning reinforces 32 rows alignment, each processing element can safely write into their assigned blocks. However, writing out the active masks poses a challenge because a single 4-byte mask encodes the active states of all 32 rows in a partition. This causes the processing elements to potentially share a single 64-byte block when writing the masks out to memory. To alleviate the contention, FlexGraph keep N active 64-byte blocks locally assigned to either one of the N processing elements. It assigns an ownership bit mask to each block to track their reference count and the processing element assigned to them. It also tracks the current address value assigned to each block. When a processing element is ready to write its 4-byte active mask, it passes the mask address and value to the LSU. The LSU goes to last active block the processing element previous wrote to and check if the block address matches. If so, it simply adds the new content to the existing block. If there is no match, it clears its ownership bit and flushes the block to memory if the ownership mask goes to zero. Then it looks up the other active block if anyone already has the address to use it, otherwise if it acquires the last unused block. Figure \ref{fig:write_masks_synchronization} shows a simplified illustration of the scheme with two processing elements. We don't need to keep more than N blocks in local storage for this scheme to work because the block address references are always incremental and never regress, meaning that there is always going to be free block available to use. It is important to also point out that for this scheme to work, the operation has to be atomic. FlexGraph has a single communication channel between all processing elements and the LSU which does some round robin arbitration and blocks the write mask request until it is committed. We also investigated an alternative solution to avoid this synchronization, which is simply increasing the partition size to 16 * 32 rows, allowing the aggregated active masks to occupy a full 64-byte block. They were two major issues with this approach. Firstly, the total size of all resident partitions will be too large if we support multiple processing elements. Secondly, it will worsen the workload imbalance that is already present with 32 rows.

This section describes some optimizations we added to FlexGraph to improve its performance.

\subsection{Optimizing Memory Accesses via Stream Buffers}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.4\textwidth]{figures/stream_buffers}
\caption{Stream Buffers}
\label{fig:stream_buffers}
\end{figure}

As stated before, FlexGraph's LSU transfer granularity is 64 bytes, which is much larger than the 4 bytes of elements accessed by the processing elements. To save on bandwidth, the LSU uses stream buffers that fetch a 64-byte blocks of data but extracts 4-byte elements at the time. It is implemented using a fifo structure in the back-end where 64-byte responses from QPI go
 into. In the front-end, there is a temporary 64-byte block that is extracted from the fifo on demand to deliver 4-byte element at the time. We employ a shift register to extract the 4-byte element to pass it to the processing element. Figure \ref{fig:stream_buffers} shows an illustration of the stream buffer concept.

\subsection{Caching Vertex Values and Masks}

Another optimization we implemented in FlexGraph is the caching of vertex values and masks. Looking at the SPMV speudo-code in listing 1, we can observe in lines 5 and 6 that there is a random memory access to buffers \textit{x\_values} and \textit{x\_activemask}. FlexGraph employs a stream buffer like concept to consume 64-byte vertex data once they arrive inside the processing element to each iteration loop. However, the same block can be referenced by another processing element and sharing them inside the cache structure can save unnecessary memory traffic. We implemented two small fully associative caches with first-in-First-out replacement to hold active vertex data during execution. The caches are implemented inside the LSU which is responsible for managing them. When a vertex request arrives from a processing element, the LSU first looks up the cache if the block is already present and return it. If the block is not there, it sends the request to QPI. QPI output interface support a 14-bit metadata field that is used to identify the block once it is returned. We use that field to store the index of the block such that we can compute the cache tag when it arrives. Because the cache is small, we implemented a one cycle tag lookup for the LSU to known if the block is present and accessible.

\subsection{Executing Non-blocking Memory writes}

FlexGraph memory writes are all non-blocking, this applies to both the output values and active masks. This allows the processing elements to push their write request and resume execution while the LSU is processing them. Upon write responses from the QPI channel, the LSU internally keeps track of the count of outstanding requested to ensure that all writes have completed. The QPI interface exposes two write response ports by which the memory replies could be sent. This allows servicing two responses simultaneously. The LSU implements a mechanism to process both channels simultaneously when they are active. At completion time when all processing elements are done executing the current run, the controller waits for all outstanding write requests to complete before asserting the \textit{done} signal.

>>>>

\begin{figure}[htbp] 
\begin{equation}
\#Edges = A^{T} \times Identity
\end{equation}
\begin{equation}
\#Edges = 
\left(\begin{array}{ccc} 0 & 0 & 0 \\ 1 & 0 & 0 \\ 1 & 1 & 0 \end{array}\right) \times
\left(\begin{array}{c} 1 \\ 1 \\ 1 \end{array}\right) 
= \left(\begin{array}{c} 0 \\ 1 \\ 2 \end{array}\right) 
\end{equation}
\caption{Incoming Edges Algorithm}
\label{fig:sample_algorithm}
\end{figure}

\begin{lstlisting}[
	emph={Assign, Process, Apply}, emphstyle=\color{blue}, 
	numbers=none, 
	escapechar=\%, 
	caption=Graph Vertex Processing Model,
	label={lst:vertex_model}]
GraphProgram(%\textit{G}%, %\textit{P}%, %\textit{N}%) :
  for_each %\textit{N}% :
    x := Assign(%\textit{P}%)
  	y := Process(%\textit{G}%,x)
  	%\textit{P}% := Apply(%\textit{P}%,y)

@@

Listing \ref{lst:vertex_model} shows the three stages of a generalized graph vertex processing model - \textit{Assign}, \textit{Process} and \textit{Apply}, given a graph \textit{G}, some vertex properties \textit{P} and an iteration count \textit{N}. The \textit{Assign} stage generates an input vector \textit{x} using the vertex properties \textit{P}. The \textit{Process} stage applies the input vector \textit{x} to the graph \textit{G} and generates an output vector \textit{y}. The\textit{ Apply} gather the resulting output vector to update the vertex properties for the next iteration. The graph program executes the three stages for several iterations until it converges. A direct mapping of our incoming edges algorithm to this processing model is also provided in Listing 1, where the identity vector is used for the vector properties. The \textit{Assign} and \textit{Apply} stages of this program are simple identity operators, while the \textit{Process} stage perform a matrix-vector multiplication. The input graph can be represented as a compressed sparse matrix to only store the non-zero entries of the graph. This data structure enables accelerated computation, often using a Sparse Matrix-Sparse Vector Multiplication (SpMSpV) kernel to target multi-core CPUs \cite{GraphMat} or GPUs \cite{MapGraph}. The computation that cannot be accelerated is simply executed on the general-purpose host CPU. FlexGraph's objective is to provide a more energy efficient graph processing backend using a FPGA for hardware acceleration \cite{Catapult}.

@@

@@

\begin{figure}[htbp]
\centering
\begin{subfigure}{0.25\textwidth}
    \includegraphics[width=\textwidth]{figures/offload_model}
    \caption{Offload model}
    \label{fig:offload_model}
\end{subfigure}
\hfill
\begin{subfigure}{0.21\textwidth}
    \includegraphics[width=\textwidth]{figures/collaborative_model}
    \caption{Collaborative model}
    \label{fig:collaborative_model}
\end{subfigure}
\caption{Shared Memory Compute Models}
\label{fig:compute_model}
\end{figure}

@@

\begin{figure}[htbp]
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=\textwidth]{figures/csc_format}
    \vspace{0pt}
\end{subfigure}
\begin{subfigure}{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/dcsc_format}
\end{subfigure}
\caption{Matrix formats comparison}
\label{fig:DCSC_matrix_format}
\end{figure}

FlexGraph uses the Doubly Compressed Sparse Column (DCSC) \cite{DCSC} format as underlying graph data structure. The format allows minimal traversal into the sparse matrix structure, saving necessary memory bandwidth when fetching empty columns. Figure \ref{fig:DCSC_matrix_format} shows the memory layout for a sample matrix A = \{(0,5,a), (0,7,b), (6,3,c), (7,1,d)\} with four non-zero edges (src, dst, weight) encoded using DCSC versus the conventional CSC \cite{CSC} format. The traversal over CSC requires accessing all consecutive column ranges \textit{(cols[i+2] - cols[i])} even though most of the distances are empty. DCSC saves on bandwidth by using an additional indirection  buffer inside its structure to only encode non-zero columns. It is important to note that this indirection can introduce additional storage overhead and access latency if the matrix is too small or not sparse enough. However, Graph Analytics datasets are generally very large and hyper-sparse which eliminate the DCSC format overhead. In addition to a sparse matrix, FlexGraph also uses a sparse vector to provide property data as well as edge selection during computation. A bitmask is used to encode the non-zero entries in the input vector.

@@

\begin{lstlisting}[
	linebackgroundcolor={
	\ifnum\value{lstnumber}=6\color{red!10}\fi
	\ifnum\value{lstnumber}=8\color{red!10}\fi
	\ifnum\value{lstnumber}=5\color{orange!10}\fi
	\ifnum\value{lstnumber}=10\color{orange!10}\fi}, 
    escapechar=\%, 
    emph={col_ind, col_ptr, row_ind, row_ptr, a_val}, emphstyle=\color{blue}, 
    caption=Pseudo-code for SpMSpV kernel,
    label={lst:spmv_code}]
def SpMSpV_kernel(a, x):
  y_values[] = {0}
  y_bitmask[] = {false}
  for_each i in a.col_ptr
    (col_ind, row_ptr) := a.col_data[i]
    x_active = x.bitmask[col_ind]          
    if (x_active):                            
      x_val = x.values[col_ind]              
      for_each j in row_ptr
        (row_ind, a_val) := a.row_data[j]
        y_values[row_ind] += a_val * x_val
        y_bitmask[row_ind] = true          
  return (y_values, y_bitmask)
\end{lstlisting}

FlexGraph micro-architecture implements a Sparse Matrix-Sparse Vertex Multiplication (SpMSpV) kernel in FPGA. This kernel diverges from conventional Sparse Matrix-Vector Multiplication (SpMV) implementations in two ways - firstly, it uses a sparse data structure as input and output vector during computation - secondly, it uses a doubly-compressed sparse matrix (DCSC) format with additional memory indirection buffer when accessing the matrix columns. These two properties present unique performance challenges when designing the accelerator. Listing \ref{lst:spmv_code} shows the pseudo-code of the SpMSpV kernel. Given the start/end addresses (col\_start, col\_end) into the matrix columns buffer (coldata), for The program iterates through each column entry and fetches the corresponding column index (col) and rows buffer address (row\_start, row\_end). It then checks if the current column index (col) is active using the vertex bitmask, only then it proceeds with rows traversal loop where the matrix row\_data is accessed to retrieve corresponding row index and matrix value for computation. A Multiply-Accumulate (MAC) operation is then performed on the matrix and vertex data and output bitmask is updated. Lines 6 and 8 show the code region where external memory is randomly accessed to obtain the vertex data and active state. Lines 5 and 10 show the code region where external memory is accessed semi-random because only the first access cannot be predicted and after that the address is simply incremented. FlexGraph's architecture attempts to address some of those performance hogs using several optimizations detailed in section 4.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\textwidth]{figures/microarchitecture}
\caption{FlexGraph Microarchitecture}
\label{fig:microarchitecture}
\end{figure}

Figure \ref{fig:microarchitecture} illustrates the microarchitecture of FlexGraph accelerator. Externally, FlexGraph input and output ports implement an Accelerator Functional Unit (AFU) interface defined by Intel's Accelerator Abstraction Layer (AAL) \cite{Intel-FPGA}. AFUs implementing the interface are able to bind with the FPGA's board support package (BSP) and seamlessly communicate with the AAL runtime on the host processor. The interface exposes four signals \textit{qpi, ctx, start, done}, where the \textit{qpi} bus handles shared-memory communications, \textit{ctx} is used for passing custom input context to the accelerator, in our case we pass down all data structure layout information needed to access memory, \textit{start} and \textit{done} are the control knobs for starting and ending execution. At a high-level, Flexgraph's
microarchitecture consists for a six building blocks: The main controller, the DCSC matrix walkers, the processing elements, the stream writers, the vertex caches and the load/store unit (LSU).

\subsection{Main Controller}

Upon reception of the \textit{start} signal from the host processor, the main controller is responsible for scheduling matrix partitions for execution on each processing element. It is comprised of two execution threads: A request thread that fetches partition data from the load/store unit into a local buffer, a dispatch thread that submit received partition data to individual matrix walker engines for processing. After all matrix partitions have been processed, the controller sends performance counters back to the host processor and activates the \textit{done} signal.

\subsection{DCSC Matrix Walkers}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\textwidth]{figures/micro_walker}
\caption{DCSC Walker}
\label{fig:micro_walker}
\end{figure}

The DCSC Matrix Walker is a custom data structure traversal engine for the DCSC Matrix. Figure \ref{fig:micro_walker} shows an overview of the engine microarchitecture. It is implemented as a four-stage state machine where each stage accesses a portion of the matrix layout as reflected in the pseudo-code in Listing \ref{lst:spmv_code}. In the first stage, the engine fetches column data (\textit{col\_ind, row\_ptr}) from memory given the partition address. In the second stage, the engine fetches the vertex active mask from memory given the \textit{col\_ind} value obtained in the previous stage. In the third stage, the engine checks if the current property is active using the mask obtained in the previous stage, if so passes the associated column data to the next stage. In the fourth stage, the engine fetches the vertex data and the matrix row data (\textit{row\_ind, a\_val}) from memory for each active column data and send those values to the connected processing element where the actual computation takes place. Each fetch unit is implemented as a Stream Reader engine to handle coarse-grain memory access for the associated data. It consists of a request module for sending memory read access command, a fifo buffer for receiving requested blocks and a temporary block buffer for accessing sub-block elements. The vertex data (values, masks) are fetched from dedicated vertex caches to enable reuse by the walker engines. We use a crossbar to multiplex outgoing requests from the walker engines to the LSU and vertex caches.

\subsection{Processing Elements}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\textwidth]{figures/micro_pe}
\caption{Processing Element}
\label{fig:micro_pe}
\end{figure}

Figure \ref{fig:micro_pe} shows an overview of a processing element microarchitecture. It main objective is to perform a floating-point multiplication on the incoming vertex and matrix values from the connected walker engine and accumulate the result until the whole partition has been processed. It is comprised of eight parallel pipelined Multiply-Accumulate (MAC) units attached to a coalescing output module. Eight is the maximum number of matrix row data that can be stored in a single cache block. Each MAC lane consists of pipelined floating-point multiply module, a pipelined floating-point Add unit, a Read-After-Write (RAW) control unit that stalls the pipeline if the current \textit{row\_ind} is pending addition. We build a bitmask given the \textit{row\_ind} to track outstanding operations in addition units. The accumulated result is stored into a local buffer indexed by the \textit{row\_ind} value. Each processing element also computes the aggregate active rows mask for the given partition. At the end of the partition processing, the coalescing unit merges output values and masks according to their associated block address and sends the result to the Stream Writers.

\subsection{Stream Writers}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\textwidth]{figures/micro_writer}
\caption{Stream Writer}
\label{fig:micro_writer}
\end{figure}

The Stream Writers are coalescing write buffers that batches write requests to memory until entire blocks are updated before they are submitted. We used two stream writers, one for the storing vertex output and one for storing the output mask. Figure \ref{fig:micro_writer} shows an overview of a stream writer microarchitecture. Internally, a stream writer has an architecture similar to a cache but has been augmented with block ownership information and the stored value is merged, not overwritten. It consists of four main modules: A data store, A tag store, and lookup unit and an eviction unit. The data store holds all the blocks sent to the device. The tag store keeps the meta-data associated with each block in the data store. A tag consists of the block address and an owners list. The owners list is implemented as a bitmask that tracks the processing elements currently using a given block. In a typical operation where the writer receives a block write request, the lookup unit is first invoked to find an existing entry for the corresponding block in the tag store. If there is a match and the block is already owned, the incoming value is simply merged with existing one. If a match was found, but the block is not owned, the previosuly owned block is released and this new one is acquired before data update. If no match is found, the device allocates a new block for the request and evicts existing data to memory if dirty. The writer also support an explicit flush command to the eviction of all dirty blocks to memory when all processing is completed.

\subsection{Vertex Caches}

FlexGraph uses two direct-mapped vertex caches to keep the vertex values and mask in the accelerator for reuse by the matrix walker engines. Their source port is connected to the load/store unit via an arbiter that multiplexes outgoing read requests with the controller and the matrix walkers. It has a typical cache architecture with an input buffer for tracking outstanding requests.

@@
